\documentclass{article}
\usepackage{parskip}

\begin{document}
In this chapter, we will examine some undefined behaviors and their mitigations if any.

\subsection{Goto}
The most glaring case of undefined behavior is the ``infinite jump'', caused by a two tables jumping back and forth.
\begin{verbatim}
(pip
  (table error1 exact
    (actions
      (set (bitfield key (int i64 0) (int i64 64)))
      (match)
    ) ; actions
    (rules
      (rule (miss)
        (actions (goto (error2)))
      )
    )
  )
  (table error2 exact
    (actions
      (set (bitfield key (int i64 0) (int i64 64)))
      (match)
    ) ; actions
    (rules
      (rule (miss)
        (actions (goto (error1)))
      )
    )
  )
)
\end{verbatim}
This is an invalid program in Pip. All tables have sequential integer identifiers, and in order for a \textbf{goto} action to be valid, the table being jumped to must have a higher identifier than the current table. In other words, the table being jumped to must come later in the program.

\section{Terminators}
The actions \textbf{drop}, \textbf{output}, and \textbf{match} are \textit{terminator} actions, meaning they can only come as the last action in an action list. All action lists end in a terminator. Thus the following undefined-behavior-invoking program is ill-formed.
\begin{verbatim}
(pip
  (table error exact
    (actions
      (set (bitfield key (int i64 0) (int i64 64)))
      (match)
    ) ; actions
    (rules
      (rule (miss)
        (actions (output (reserved_port controller)) 
                 (output (reserved_port in_port)))
      )
    )
  )
)
\end{verbatim}

The \textbf{match} terminator action is slightly different than the others. All exact-match tables \textit{must} begin their matching phase, even if no rules are defined. The \textbf{match} action \textit{must} appear at the end of the table's prep-list, or action list that sets the key. Thus, the following program is ill-formed.
\begin{verbatim}
(pip
  (table error exact
    (actions
      (set (bitfield key (int i64 0) (int i64 64)))
    ) ; actions
    (rules
      (rule (miss)
        (actions (output (reserved_port controller)))
      )
    )
  )
)
\end{verbatim}
As is a program that matches outside of the prep-list.
\begin{verbatim}
(pip
  (table error exact
    (actions
      (set (bitfield key (int i64 0) (int i64 64)))
      (match)
    ) ; actions
    (rules
      (rule (miss)
        (actions (clear) (match))
      )
    )
  )
)
\end{verbatim}

\subsection{Keyless Programs}
An exact match table \textit{must} set its key register in the prep-list. The following program is ill-formed.
\begin{verbatim}
(pip
  (table error exact
    (actions
      (match)
    ) ; actions
    (rules
      (rule (miss)
        (actions (drop))
      )
    )
  )
)
\end{verbatim}
\end{document}
